---
title: "Bellabeat Case Study"
author: "YH"
date: "July 2, 2025"
output: html_document
---

```{r setup, include=FALSE}
# This chunk loads all necessary libraries for the analysis.
# 'include=FALSE' means this code chunk will run, but won't be shown in the final report.
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)  # For data manipulation and visualization (includes ggplot2)
library(lubridate)  # For working with dates
library(janitor)    # For cleaning column names
```

# 1. Ask

In this section, I will define the business task and the key questions that this analysis aims to answer. The goal is to understand how consumers use their non-Bellabeat smart devices and translate these insights into actionable marketing recommendations for Bellabeat.

### 1.1 Guiding Questions

1.  **What are some trends in smart device usage?**
2.  **How could these trends apply to Bellabeat customers?**
3.  **How could these trends help influence Bellabeat marketing strategy?**
  
### 1.2 Business Task

**To analyze the FitBit Fitness Tracker dataset to identify usage trends and provide data-driven recommendations for Bellabeat's marketing strategy.**

### 1.3 Key Stakeholders

*   **Urška Sršen:** Cofounder and Chief Creative Officer of Bellabeat.
*   **Sando Mur:** Cofounder and mathematician.
*   **Bellabeat Marketing Analytics Team:** The team you are a part of.


# 2. Prepare

This phase involves sourcing, loading, and exploring the data. The data used is the **FitBit Fitness Tracker Data**, a public dataset available on Kaggle, which contains information from thirty Fitbit users over a 31-day period.

### 2.1 Loading Data

The following code loads the relevant CSV files into R dataframes.

```{r load_data}
# Use the path to where you saved the files on your computer
daily_activity <- read_csv("FitBit_Fitness_Tracker_Data/mturkfitbit_export_4.12.16-5.12.16/Fitabase Data 4.12.16-5.12.16/dailyActivity_merged.csv")
daily_sleep <- read_csv("FitBit_Fitness_Tracker_Data/mturkfitbit_export_4.12.16-5.12.16/Fitabase Data 4.12.16-5.12.16/sleepDay_merged.csv")
daily_calories <- read_csv("FitBit_Fitness_Tracker_Data/mturkfitbit_export_4.12.16-5.12.16/Fitabase Data 4.12.16-5.12.16/dailyCalories_merged.csv")
weight <- read_csv("FitBit_Fitness_Tracker_Data/mturkfitbit_export_4.12.16-5.12.16/Fitabase Data 4.12.16-5.12.16/weightLogInfo_merged.csv")
```

### 2.2 Initial Data Exploration

Let's preview the structure and contents of each dataset to understand what we're working with.

```{r preview_data}
head(daily_activity)
head(daily_sleep)
str(daily_activity)
```

### 2.3 Data Limitations (ROCCC Assessment)

*   **Reliable:** The data comes from only 30 users, which is not a representative sample of the broader population. Findings should be treated as directional.
*   **Original:** This is third-party data (from a Kaggle user named Mobius), not from Bellabeat's own user base.
*   **Comprehensive:** The data includes activity, sleep, and some weight logs, but lacks information on stress, nutrition, and menstrual cycles, which are key features for Bellabeat.
*   **Current:** The data was collected in 2016. User behaviors and device technology have changed significantly since then, which is a major limitation.
*   **Cited:** The data is cited and publicly available, but its collection method is not fully detailed.


# 3. Process

In this phase, the data is cleaned, transformed, and organized for analysis. Key steps include standardizing column names, formatting dates, checking for duplicates, and merging datasets.

### 3.1 Standardizing Column Names & Formatting Dates

Let's clean the column names to be consistent (snake_case) and ensure the date columns are in the correct format.

```{r clean_and_format}
# Clean names
daily_activity <- clean_names(daily_activity)
daily_sleep <- clean_names(daily_sleep)

# Format dates
daily_activity$activity_date <- as.Date(daily_activity$activity_date, format="%m/%d/%Y")
daily_sleep$sleep_day <- as.Date(daily_sleep$sleep_day, format="%m/%d/%Y")
```

### 3.2 Checking for Duplicates and NAs

Now we check for and remove any duplicate entries and inspect for missing values.

```{r check_duplicates_na}
# Count duplicates
sum(duplicated(daily_activity))
sum(duplicated(daily_sleep))

# Remove duplicates
daily_activity <- daily_activity %>% distinct()
daily_sleep <- daily_sleep %>% distinct()

# Check for NAs
colSums(is.na(daily_activity))
colSums(is.na(daily_sleep))
```

### 3.3 Merging Datasets

To analyze relationships between activity and sleep, we will merge the two main datasets into one.

```{r merge_data}
# Rename the date columns to be consistent before merging
daily_activity <- daily_activity %>% rename(date = activity_date)
daily_sleep <- daily_sleep %>% rename(date = sleep_day)

# Merge the two dataframes
activity_sleep_merged <- merge(daily_activity, daily_sleep, by=c("id", "date"))
head(activity_sleep_merged)
```

# 4. Analyze

This phase focuses on exploring the data to pull out key insights. We will calculate summary statistics and look for relationships between different variables.

### 4.1 Summary Statistics

Let's get a high-level overview of user habits.

```{r summary_stats}
activity_sleep_merged %>%
  select(total_steps,
         total_distance,
         sedentary_minutes,
         total_minutes_asleep) %>%
  summary()
```

### 4.2 Exploring Relationships

Now, let's dive deeper. How does activity relate to calories burned? How does time in bed relate to actual sleep?

```{r analyze_relationships}
# Relationship between steps and calories
ggplot(data = activity_sleep_merged, aes(x = total_steps, y = calories)) +
  geom_point(color = "purple") +
  geom_smooth(method = "loess") +
  labs(title = "Total Steps vs. Calories Burned",
       x = "Total Steps",
       y = "Calories Burned")

# Relationship between time in bed and time asleep
ggplot(data = activity_sleep_merged, aes(x = total_time_in_bed, y = total_minutes_asleep)) +
  geom_point(color = "orange") +
  geom_abline(slope = 1, intercept = 0, color="blue") + # Adds a 1:1 line
  labs(title = "Time in Bed vs. Time Asleep",
       x = "Total Time in Bed (Minutes)",
       y = "Total Minutes Asleep")
```

# 5. Share

This phase is about communicating the story found in the data. Here are three key findings supported by visualizations.

### Finding 1: Users are Largely Sedentary

Despite some users being very active, the average user spends the vast majority of their day (about 81%) being sedentary. This presents a major opportunity for features that encourage movement.

```{r visualize_activity_levels}
# Calculate average minutes for each activity level
avg_minutes <- activity_sleep_merged %>%
  summarise(
    avg_sedentary = mean(sedentary_minutes),
    avg_lightly_active = mean(lightly_active_minutes),
    avg_fairly_active = mean(fairly_active_minutes),
    avg_very_active = mean(very_active_minutes)
  ) %>%
  gather(key = "activity_level", value = "average_minutes")

# Create the bar chart
ggplot(data = avg_minutes, aes(x = activity_level, y = average_minutes, fill = activity_level)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Daily Time Spent in Each Activity Level",
       x = "Activity Level",
       y = "Average Minutes per Day") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Finding 2: Inefficient Sleep is Common

The analysis of time in bed versus time asleep shows that many data points fall below the blue line, which represents 100% sleep efficiency. This gap indicates that users spend time in bed awake, suggesting an opportunity to help users improve their sleep quality.

*(The visualization for this was already created in the Analyze phase. Here, you would add more text to explain its business implications.)*

### Finding 3: Activity Patterns Vary by Day

By creating a "day of the week" column, we can see how activity levels change. For instance, users might be more active on weekends.

```{r analyze_day_of_week}
activity_sleep_merged$day_of_week <- wday(activity_sleep_merged$date, label = TRUE)

ggplot(data = activity_sleep_merged, aes(x = day_of_week, y = total_steps, fill = day_of_week)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Steps by Day of the Week",
       x = "Day of the Week",
       y = "Total Steps")
```

# 6. Act

Based on the analysis, here are three high-level recommendations for Bellabeat's marketing strategy.

### Recommendation 1: Market "Move Reminders" to Address Sedentary Behavior

*   **The Insight:** Users are sedentary for over 16 hours a day on average.
*   **The Recommendation:** Launch a marketing campaign focused on a new or existing feature in the Bellabeat app and devices (like Leaf and Time) that provides smart notifications to encourage users to take short walks or stretch. Marketing messaging should highlight how small movements can break up the day and improve overall wellness.

### Recommendation 2: Focus on Sleep Quality, Not Just Duration

*   **The Insight:** Users often spend significant time in bed without being asleep, indicating poor sleep efficiency.
*   **The Recommendation:** Enhance the Bellabeat app with content about sleep hygiene (e.g., creating a bedtime routine, reducing screen time). Marketing should position Bellabeat not just as a sleep tracker, but as a "sleep coach" that provides actionable advice to improve the quality of sleep.

### Recommendation 3: Create Targeted Weekend and Weekday Content

*   **The Insight:** Activity patterns, particularly total steps, can vary across the week.
*   **The Recommendation:** Develop targeted content and app notifications for weekdays versus weekends. For example, weekday content could focus on "office wellness" and fitting in activity during a busy schedule, while weekend content could promote longer, more intensive activities. This personalized approach can make marketing feel more relevant to the user's daily life.
