---
title: "Bellabeat Case Study: A Strategic Analysis"
author: "YH"
date: "July 17, 2025"
output: html_document
---

```{r setup, include=FALSE}
# This chunk loads all necessary libraries for the analysis.
# 'include=FALSE' means this code chunk will run, but won't be shown in the final report.
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(janitor)
```

# 1. Ask

This section defines the business objective and the key questions driving this analysis. The goal is to translate insights from non-Bellabeat smart device data into actionable marketing strategies for Bellabeat.

### 1.1 Guiding Questions

1.  **What are the key trends and user behaviors in smart device usage?**
2.  **How could these trends apply to Bellabeat's user base?**
3.  **How can these insights influence Bellabeat's marketing strategy?**
  
### 1.2 Project Objective (Refinement 1)

**To analyze third-party fitness tracker data to deliver data-driven recommendations that guide Bellabeat's future marketing strategy.**

### 1.3 Key Stakeholders

*   **Urška Sršen**: Cofounder and Chief Creative Officer
*   **Sando Mur**: Cofounder and Mathematician
*   **Bellabeat Marketing Analytics Team**

# 2. Prepare

This phase involves sourcing, loading, and assessing the data. The data is the **FitBit Fitness Tracker Data** from Kaggle, containing data from thirty Fitbit users over a 31-day period.

### 2.1 Loading Data

```{r load_data}
daily_activity <- read_csv("FitBit_Fitness_Tracker_Data/mturkfitbit_export_4.12.16-5.12.16/Fitabase Data 4.12.16-5.12.16/dailyActivity_merged.csv")
daily_sleep <- read_csv("FitBit_Fitness_Tracker_Data/mturkfitbit_export_4.12.16-5.12.16/Fitabase Data 4.12.16-5.12.16/sleepDay_merged.csv")
```

### 2.2 Quantifying the Sample (Refinement 2)
Before proceeding, let's confirm the number of unique participants in our datasets. This is a critical first step for understanding the scope of the data.

```{r count_participants}
n_participants_activity <- n_distinct(daily_activity$Id)
n_participants_sleep <- n_distinct(daily_sleep$Id)

cat("Number of unique participants in activity data:", n_participants_activity, "\n")
cat("Number of unique participants in sleep data:", n_participants_sleep, "\n")
```
*The analysis is based on approximately 30 users, which is a small sample size. This will be noted in the limitations.*

### 2.3 Data Limitations (ROCCC Assessment)

*   **Reliable:** **Low.** The data comes from only ~30 users, which is not a representative sample. Findings are directional.
*   **Original:** **No.** This is third-party data from Kaggle.
*   **Comprehensive:** **Medium.** Lacks key Bellabeat metrics like stress, nutrition, and menstrual cycles.
*   **Current:** **Low.** Data is from 2016. User behaviors have likely changed.
*   **Cited:** **Yes.** Data is publicly available but lacks detailed collection methodology.

# 3. Process

This phase cleans and transforms the data for analysis. This includes standardizing column names, formatting dates, removing duplicates, merging datasets, and engineering new features.

### 3.1 Standardizing Column Names & Formatting Dates

```{r clean_and_format}
daily_activity <- clean_names(daily_activity)
daily_sleep <- clean_names(daily_sleep)

daily_activity$activity_date <- as.Date(daily_activity$activity_date, format="%m/%d/%Y")
daily_sleep$sleep_day <- as.Date(daily_sleep$sleep_day, format="%m/%d/%Y")
```

### 3.2 Checking for Duplicates and NAs

```{r check_duplicates_na}
daily_activity <- daily_activity %>% distinct()
daily_sleep <- daily_sleep %>% distinct()
```

### 3.3 Merging Datasets & Feature Engineering (Refinement 3)

We merge the two main datasets to analyze relationships between activity and sleep. We will also engineer a `day_of_week` feature here to centralize all data processing steps.

```{r merge_and_feature_engineer}
daily_activity <- daily_activity %>% rename(date = activity_date)
daily_sleep <- daily_sleep %>% rename(date = sleep_day)

activity_sleep_merged <- merge(daily_activity, daily_sleep, by=c("id", "date"))

# Add 'day_of_week' feature
activity_sleep_merged$day_of_week <- wday(activity_sleep_merged$date, label = TRUE)

head(activity_sleep_merged)
```

# 4. Analyze

This phase explores the processed data to uncover insights through summary statistics, relationship analysis, and user segmentation.

### 4.1 Summary Statistics

```{r summary_stats}
activity_sleep_merged %>%
  select(total_steps, total_distance, sedentary_minutes, total_minutes_asleep) %>%
  summary()
```

### 4.2 Exploring Relationships

```{r analyze_relationships}
# Relationship between steps and calories
ggplot(data = activity_sleep_merged, aes(x = total_steps, y = calories)) +
  geom_point(color = "purple", alpha = 0.5) +
  geom_smooth(method = "loess") +
  labs(title = "Total Steps vs. Calories Burned", x = "Total Steps", y = "Calories Burned")

# Relationship between time in bed and time asleep
ggplot(data = activity_sleep_merged, aes(x = total_time_in_bed, y = total_minutes_asleep)) +
  geom_point(color = "orange", alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0, color="blue") +
  labs(title = "Time in Bed vs. Time Asleep", x = "Total Time in Bed (Minutes)", y = "Total Minutes Asleep")
```

### 4.3 User Segmentation by Activity Level (Refinement 4)
To better understand our users, we can segment them based on their average daily steps. This provides a more nuanced view than looking at the population as a whole. (Using CDC guidelines: <5,000=Sedentary, 5,000-7,499=Lightly Active, >7,500=Active).

```{r segment_users}
user_avg_steps <- activity_sleep_merged %>%
  group_by(id) %>%
  summarise(avg_steps = mean(total_steps)) %>%
  mutate(user_type = case_when(
    avg_steps < 5000 ~ "Sedentary",
    avg_steps >= 5000 & avg_steps < 7500 ~ "Lightly Active",
    avg_steps >= 7500 ~ "Active"
  ))

# Show the percentage of users in each segment
user_segment_summary <- user_avg_steps %>%
  group_by(user_type) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(user_segment_summary, aes(x = "", y = percentage, fill = user_type)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  geom_text(aes(label = paste0(round(percentage), "%")), position = position_stack(vjust = 0.5)) +
  labs(title = "User Segmentation by Activity Level", fill = "User Type", x = NULL, y = NULL) +
  theme_void()
```

# 5. Share

This section communicates the key findings from the analysis, supported by clear visualizations.

### Finding 1: Users are Largely Sedentary

The average user spends the vast majority of their day (about 81%) being sedentary. This presents a major opportunity for features that encourage movement.

```{r visualize_activity_levels_pie, fig.width=8, fig.height=6}
# Calculate average minutes for each activity level
avg_minutes <- activity_sleep_merged %>%
  summarise(
    Sedentary = mean(sedentary_minutes),
    `Lightly Active` = mean(lightly_active_minutes),
    `Fairly Active` = mean(fairly_active_minutes),
    `Very Active` = mean(very_active_minutes)
  ) %>%
  gather(key = "activity_level", value = "average_minutes") %>%
  mutate(activity_level = factor(activity_level, levels = c("Very Active", "Fairly Active", "Lightly Active", "Sedentary")))

# Create the pie chart (Refinement 5)
ggplot(data = avg_minutes, aes(x = "", y = average_minutes, fill = activity_level)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  geom_text(aes(label = paste0(round(average_minutes / sum(average_minutes) * 100), "%")),
            position = position_stack(vjust = 0.5)) +
  labs(title = "Proportion of Day Spent in Each Activity Level",
       subtitle = "Sedentary behavior dominates the average user's day",
       fill = "Activity Level",
       x = NULL, y = NULL) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5, size=16),
        plot.subtitle = element_text(hjust = 0.5, size=12))
```

### Finding 2: Inefficient Sleep is Common
The analysis of time in bed versus time asleep shows that many data points fall below the 1:1 line, indicating users spend time in bed awake. This suggests an opportunity to help users improve their sleep quality.

### Finding 3: Activity Patterns Vary by Day
By analyzing total steps by the day of the week, we can see potential patterns, such as higher activity on weekends.

```{r analyze_day_of_week}
ggplot(data = activity_sleep_merged, aes(x = day_of_week, y = total_steps, fill = day_of_week)) +
  geom_boxplot() + # A boxplot is better to show distribution
  labs(title = "Distribution of Steps by Day of the Week",
       x = "Day of the Week",
       y = "Total Steps") +
  theme(legend.position = "none")
```

# 6. Act

Based on the analysis, here are three high-level recommendations for Bellabeat's marketing strategy.

### Recommendation 1: Launch a Targeted "Mindful Movement" Campaign (Refinement 6)

*   **The Insight:** Users are sedentary for over 16 hours a day on average, and our segmentation shows a significant portion of users fall into the "Sedentary" category.
*   **The Recommendation:** Launch a targeted marketing campaign and in-app feature called "Mindful Movement." This feature would send smart, encouraging notifications to users in the "Sedentary" segment to take short walks or stretch.
*   **Business Impact:** The campaign's success will be measured via an A/B test. **The goal is to achieve a 15% reduction in average daily sedentary minutes for the test group over a 3-month pilot.** This would directly address a core health concern and increase user engagement with the Bellabeat app and devices.

### Recommendation 2: Focus on Sleep Quality, Not Just Duration

*   **The Insight:** Users often spend significant time in bed without being asleep, indicating poor sleep efficiency.
*   **The Recommendation:** Enhance the Bellabeat app with content about sleep hygiene. Marketing should position Bellabeat not just as a sleep tracker, but as a "sleep coach" that provides actionable advice.

### Recommendation 3: Create Targeted Weekend and Weekday Content

*   **The Insight:** Activity patterns can vary across the week.
*   **The Recommendation:** Develop targeted content and app notifications for weekdays versus weekends. For example, weekday content could focus on "office wellness," while weekend content could promote longer, more intensive activities.
